Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-19 paper/ms.tex
a few words for positive definiteness

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/transit-ml.pdf}
    \caption{\emph{(top)} A month-long segment of \kepler\ light curve for
    KIC~1430163 with a synthetic transit model injected (black points) and the
    maximum likelihood model for the stellar variability and the transit model
    (blue line).
    \emph{(bottom)} The maximum likelihood ``de-trendeding'' of the data in
    the top panel.
    In this panel, the maximum likelihood model for the stellar variability
    has been subtracted to leave only the transits.
    The de-trended light curve is shown by black error bars and the maximum
    likelihood transit model is shown as a blue line.
    \figurelabel{transit-ml}}
\end{center}
\end{figure}

@@

\acknowledgments
It is a pleasure to thank
Sivaram Ambikasaran,
Megan Bedell,
Will Farr,
Sam Grunblatt,
David W.\ Hogg, and
Dan Huber
for helpful contributions to the ideas and code presented here.

@@

\section{Ensuring positive definiteness: Sturm's theorem}

>>>>

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/transit-ml.pdf}
    \caption{\emph{(top)} A month-long segment of \kepler\ light curve for
    KIC~1430163 with a synthetic transit model injected (black points) and the
    maximum likelihood model for the stellar variability
    (blue line).
    \emph{(bottom)} The maximum likelihood ``de-trending'' of the data in
    the top panel.
    In this panel, the maximum likelihood model for the stellar variability
    has been subtracted to leave only the transits.
    The de-trended light curve is shown by black error bars and the maximum
    likelihood transit model is shown as a blue line.
    \figurelabel{transit-ml}}
\end{center}
\end{figure}

@@

\acknowledgments
It is a pleasure to thank
Megan Bedell,
Will Farr,
Sam Grunblatt,
David W.\ Hogg, and
Dan Huber
for helpful contributions to the ideas and code presented here.

@@

\section{Ensuring positive definiteness}

For a GP kernel to be valid, it must produce a positive definite covariance
matrix for all input coordinates.
For stationary kernels, this is equivalent~--~by Bochner's theorem \citep[see
Section 4.2.1 in][]{Rasmussen:2006}~--~to requiring that the kernel be the
Fourier transform of a positive finite measure.
This means that the power spectrum of a positive definite kernel must be
positive for all frequencies.

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.
